negativeBinomial

### 2015-09-04 lengthModel/lengthModelHeterogeneous.Rmd
update

```{r}
library(ggplot2)
library(stringr)
set.seed(123)
data_length <- read.csv("~/mysis/data/lengthData.csv", skip = 4, header = TRUE)
net <- data_length$Size..m.
y <- data_length$Length..mm.
gender <- as.factor(substr(str_trim(as.character(data_length$Gender)), 0, 1))
                       ## remove whitespace from string, then truncate into 
                       ## broad gender classes, not subclasses

date <- as.numeric(data_length$Date)
date[date == 2] <- 1      ## collapse into monthly values
date[date == 3] <- 2
date[date == 4] <- 3
date <- as.factor(date)
plot(as.numeric(date), type='l', main = "Is this a data error? - yes?")
## correct date mislabeling
date <- as.numeric(data_length$Date)
date[7330:7519] <- 4
date[7559:7773] <- 4
date[date == 2] <- 1      ## collapse into monthly values
date[date == 3] <- 2
date[date == 4] <- 3
date <- as.factor(date)
plot(as.numeric(date), type='l', main = "Data looks better")
data <- data.frame(y=y, date=date, net=net, gender=gender)
X <- model.matrix(~ net + date + gender, data=data)
ggplot(data=data, aes(y)) + geom_histogram() + facet_grid(net ~ date)
```

@@

```{r, echo=FALSE}
library(rstan, verbose=FALSE)
model <- "
// regression
data {
  int<lower=1> N;             // number of observations
  int<lower=1> p;             // number of predictors
  matrix[N, p] X;             // predictor
  vector[N] y;                // response
} parameters {  
  real<lower=0> s;            // regression standard deviation
  vector[p] beta;             // mean parameters
  vector<lower=0>[p] gamma2;  // LASSO mixing parameters
  real<lower=0> lambda2;      // LASSO shrinkage parameters
} model {
  // priors:
  //s ~ cauchy(0, 20);
  lambda2 ~ gamma(1, 1);      // Prior for weak LASSO shrinkage 
  gamma2 ~ exponential(lambda2 / 2.0);
  for(i in 2:p){
    beta[i] ~ normal(0, s * sqrt(gamma2[i]));
  }

  // data model:
  y ~ normal(X * beta, s);
}
"
write(model, file = "~/mysis/stanModels/length.stan")
# sm <- stan_model("~/mysis/stanModels/length.stan")
sm <- stan("~/mysis/stanModels/length.stan", 
           data=list(N=length(y), p=dim(X)[2], y=y, X=X),
           chains=0)
```
## Fit the Model

```{r, echo=FALSE, cache=TRUE}
library(parallel)
estimate_params <- function(model, data, pars = pars, iter=5000){
  sflist <- mclapply(1:4, mc.cores = 4,
             function(i) stan(fit = model, data = data, pars = pars,
                              chains = 1, chain_id = i,
                              iter=iter,
                              refresh = -1))
  fit <- sflist2stanfit(sflist)
  return(fit)
}
# m <- sampling(sm, data = list(N=length(y), p=dim(X)[2], y=y, X=X),
#   pars = c("beta", "s"),
#   iter = 1000, chains = 4)
m <- estimate_params(sm, data = list(N=length(y), p=dim(X)[2], y=y, X=X),
                     pars = c("beta", "s", "gamma2", "lambda2"))
```

@@

>>>>

```{r}
library(ggplot2)
library(stringr)
library(mgcv)
set.seed(123)
data_length <- read.csv("~/mysis/data/lengthData.csv", skip = 4, header = TRUE)
net <- data_length$Size..m.
y <- data_length$Length..mm.
gender <- as.factor(substr(str_trim(as.character(data_length$Gender)), 0, 1))
                       ## remove whitespace from string, then truncate into 
                       ## broad gender classes, not subclasses

date <- as.numeric(data_length$Date)
date[date == 2] <- 1      ## collapse into monthly values
date[date == 3] <- 2
date[date == 4] <- 3
date <- as.factor(date)
plot(as.numeric(date), type='l', main = "Is this a data error? - yes?")
## correct date mislabeling
date <- as.numeric(data_length$Date)
date[7330:7519] <- 4
date[7559:7773] <- 4
date[date == 2] <- 1      ## collapse into monthly values
date[date == 3] <- 2
date[date == 4] <- 3
date <- as.factor(date)
plot(as.numeric(date), type='l', main = "Data looks better")
data <- data.frame(y=y, date=date, net=net, gender=gender)
X <- model.matrix(~ net + date + gender, data=data)
ggplot(data=data, aes(y)) + geom_histogram() + facet_grid(net ~ date)
idx <- attr(uniquecombs(X), "index")
```

@@

```{r, echo=FALSE}
library(rstan, verbose=FALSE)
model <- "
// regression
data {
  int<lower=1> N;             // number of observations
  int<lower=1> p;             // number of predictors
  int<lower=1> q;             // number of levels
  matrix[N, p] X;             // predictor
  vector[N] y;                // response
  int idx[N];                 // indicator variable of treatment
} parameters {  
  vector<lower=0>[q] s;       // heterogeneous regression standard deviation
  vector[p] beta;             // mean parameters
} model {
  // priors:
  //s ~ cauchy(0, 20);
  for(i in 2:p){
    beta[i] ~ normal(0, 100);
  }

  // data model:
  for(i in 1:N){
    y[i] ~ normal(X[i] * beta, s[idx[i]]);
  }
}
"
write(model, file = "~/mysis/stanModels/lengthHeterogeneous.stan")
# sm <- stan_model("~/mysis/stanModels/lengthHeterogeneous.stan")
sm <- stan("~/mysis/stanModels/lengthHeterogeneous.stan", 
           data=list(N=length(y), p=dim(X)[2], y=y, X=X, idx=idx, q=length(unique(idx))),
           chains=0)
```
## Fit the Model

```{r, echo=FALSE, cache=TRUE}
library(parallel)
estimate_params <- function(model, data, pars = pars, iter=5000){
  sflist <- mclapply(1:4, mc.cores = 4,
             function(i) stan(fit = model, data = data, pars = pars,
                              chains = 1, chain_id = i,
                              iter=iter,
                              refresh = -1))
  fit <- sflist2stanfit(sflist)
  return(fit)
}
# m <- sampling(sm, data = list(N=length(y), p=dim(X)[2], y=y, X=X, idx=idx, q=length(unique(idx))),
#   pars = c("beta", "s"),
#   iter = 1000, chains = 1)
m <- estimate_params(sm, data = list(N=length(y), p=dim(X)[2], y=y, X=X, idx=idx, q=length(unique(idx))),
                     pars = c("beta", "s"))
```

@@

```{r}
summary(lm(y ~ net + date + gender))
plot(lm(y ~ net + date + gender))
library(MASS)
summary(rlm(y ~ net + date + gender))
plot(rlm(y ~ net + date + gender))
```
